\section{Introduction}
\label{sec:intro}

Enterprise, cloud, and industrial edges now carry predominantly encrypted sessions.
Where bulk decryption is infeasible or policy-disallowed, defenders observe TLS handshakes, record sizes, inter-arrival times, and coarse metadata---signals that are legally and technically accessible but weakly aligned with traditional application-layer antivirus taxonomies.

\textbf{Gap.}
A rich literature improves \emph{encrypted traffic classification}, yet many published systems stop at an argmax label.
In contrast, SOCs, incident responders, and orchestration fabrics consume \emph{events}: artifacts with stable identifiers, timestamps, observable context, calibrated confidence, tactic-relevant semantics, and lineage to models and datasets.
Without such structure, classifiers are hard to replay after model updates, difficult to correlate across hosts, and risky to feed into automated playbooks.

\textbf{Positioning.}
We treat the learning stack as \emph{replaceable}: commodity flow export (e.g., IPFIX-style summaries~\cite{rfc7011}) and learned representations~\cite{lin2022etbert} are implementation details behind a fixed \emph{event interface}.
Our scientific question is operational: \emph{How should encrypted flow predictions be serialized so training-time validation, online inference, and offline replay agree on dispatch and final fields?}

\textbf{Approach.}
Stage~1 performs benign/malicious screening with probability calibration and a single validated threshold policy shared across train, serve, and replay.
Stage~2 assigns finer threat-behavior semantics only to flows that satisfy that gate, reflecting asymmetric compute budgets at the edge.
Each flow maps to the schema in Section~\ref{sec:design}.

\textbf{Snapshot.}
Stage~1 OOD macro-F1 is $0.9478$; Stage~2 macro-F1 is $0.7778$; $460$ replayed flows produce $460$ structured signals with $42$ Stage~2 inferences, matching dispatch counters under fixed versions.

\paragraph{Contributions.}
\begin{itemize}\itemsep1pt
  \item \textbf{Two-stage edge screening and behavior labeling} with explicitly shared calibration/threshold semantics between training and production.
  \item \textbf{Structured event-signal schema} aligning inference outputs to downstream correlation and CTI-style fields without vendor lock-in.
  \item \textbf{Replay validation} demonstrating per-flow signal cardinality and Stage~1/Stage~2 admission parity on archived traces.
\end{itemize}

Temporal feature fusion, open-set discovery, long-term drift adaptation, unseen-protocol guarantees, Edge-SLM copilots, and automated enforcement are \emph{explicitly out of scope} for the present validation (Section~\ref{sec:conclusion}).
