\section{Methodology}
\label{sec:method}

This section instantiates the design of Section~\ref{sec:design}: how features are produced, how Stage~1/Stage~2 models are trained under a single dispatch policy, and how outputs are serialized and replay-checked against that policy.

\subsection{Feature materialization}

We ingest any source honoring the Layer~1 contract (Section~\ref{sec:design}): flows keyed by five-tuple and time window, augmented with TLS metadata when present and optional per-packet summaries for future temporal modules.
Feature dictionaries are versioned; hashes over canonicalized tensors are stored in $g_i$ to detect silent skew between training and serving.

\subsection{Stage~1 screening and calibration}

The Stage~1 classifier outputs malicious probability $\hat{p}$; temperature scaling (or equivalent) maps logits to calibrated $\bar{p}$ on a held-out development split disjoint from OOD evaluation.
Class imbalance is countered via reweighting or balanced sampling; the objective is reliable ranking at the operating threshold rather than leaderboard accuracy alone.

\subsection{Threshold-consistent dispatch}

Let $\tau$ denote the malicious admission threshold selected on development replay traces.
A flow escalates iff $\bar{p}_{\mathrm{mal}} \ge \tau$.
The \emph{same} inequality governs (i)~mining supervisory tuples for Stage~2 training, (ii)~online inference, and (iii)~offline archival replay.
This alignment prevents benign label drift (e.g., exporting ``malicious'' nominal labels while dispatch logic suppresses Stage~2) and is necessary for faithful event emission.

\subsection{Stage~2 behavior model}

Stage~2 is trained exclusively on flows whose features were generated under the dispatch gate, mirroring conditional distributions at deployment.
The classifier head targets the Stage~2 taxonomy in Table~\ref{tab:label_space}; macro-F1 is emphasized due to class imbalance.

\subsection{Signal synthesis and replay}

For each flow we populate $e_i$ (Eq.~\eqref{eq:event}) including top-$k$ logits, calibrated scores, boolean dispatch flags, schema/model version tuples, and deterministic checksums of serialized inputs.
Replay replays stored features through the frozen cascade; Section~\ref{sec:experiments} verifies one signal per ingested flow and identical Stage~2 cardinality relative to admitted malicious traffic.
